---
title: "Challenge 3"
author: "Chloe Feehan"
format:
  html: 
    self-contained: true
    code-tools: true
    code-fold: true
execute:
  echo: true
  error: true
  message: false
---

# Challenge 3: Group Comparisons & Data Ethics

## Published Comparisons

In the published article ([You Know What It Is: Learning Words through Listening to Hip-Hop](https://journals.plos.org/plosone/article/file?id=10.1371/journal.pone.0028248&type=printable)), the author presents a series of comparisons about the variables which most explain an individual's familiarity with African American English (AAE).

Let us instead compare the number of artists participants reported liking for each genre. Specifically, you will determine which music genre most differentiates each of the demographic groups provided.

> Which genre had much higher average (mean or median) reported artists in one group than the other.
>
> -   Male versus Female
> -   White versus Non-White

::: callout-tip
You might find it helpful to first create a new data set with only the variables you are interested in! Look at the Music Variables in the data set description along with the demographics of interest.

Helpful functions: `select()`, `group_by()`, `summarize()`, `across()`

Other useful operations in R: `mean()`, `diff()`, `abs()`, `which.max()`
:::

```{r packages}
#| message: false
# code chunk for loading packages and importing the data
library(tidyverse)
library(here)
hiphop <- read_csv(here("Labs", "Lab 3", "hiphop.csv"))

hiphop_clean <- hiphop |>
  mutate(
    across(
      c(word, sex, ethnic, familiarity),
      as.factor
      ))

hiphop_clean <- hiphop_clean |>
  mutate(ethnic = 
    case_when(ethnic == "white"~"white",
            TRUE ~ "non-white")
  )
```

```{r}
new_hiphop <- hiphop_clean |>
  select(
    c(sex, ethnic, intl:unclassifiable)
  ) |>
  mutate(
    across(
      .col = intl:unclassifiable,
      .fns = ~as.numeric(.x)
    )
  )

```

```{r}
male_female <- new_hiphop |>
  group_by(sex) |> 
  summarize(
      across(.cols = intl:unclassifiable, 
             .fns = ~ mean(.x)
    ))

male_female <- male_female |>
  summarize(
      across(.cols = intl:unclassifiable, 
             .fns = ~ diff(.x)
    ))
  
male_female |>
  which.max()

```
Between male and females, rock had the highest average reported artists.

```{r}
white_nonwhite <- new_hiphop |>
  group_by(ethnic) |> 
  summarize(
      across(.cols = intl:unclassifiable, 
              .fns = ~mean(.x)
    ))

white_nonwhite <- white_nonwhite |>
  summarize(
    abs(
     across(.cols = intl:unclassifiable, 
              .fns = ~diff(.x)
    )))
white_nonwhite |>
 which.max()


```
Between whites and nonwhites, the international genre had the highest average reported artists.

## Study Design Critique -- Data Ethics

Myself, members of the Tidy Tuesday community, and previous 331 students have voiced concerns regarding the design and social context of this study.

You've already read the [data description (link)](http://conservancy.umn.edu/bitstream/handle/11299/116327/5/explanationAAEHiphopChesley.txt) regarding how participants were recruited for participation in this study. **Now**, you need to read additional details regarding aspects of the study in the published paper: [You Know What It Is: Learning Words through Listening to Hip-Hop](https://journals.plos.org/plosone/article/file?id=10.1371/journal.pone.0028248&type=printable).

> Based on the design of this study and its context (African American English), what are **at least two** concerns you have? Keep in mind this critique aligns with conversations regarding data ethics. Thus, your concerns need to address the racial aspects of the design of this study.

Some concerns that I have regarding the ethics of this study is that the basis of this study is testing the knowledge of an individuals AAE vocabulary based on musical preferences, social ties to African Americans, and knowledge of pop culture. However, these factors are strictly media-based and don't encapsulate the entirety of the culture. It condenses the vocabulary of AAE into stereotypes based on certain TV series. This study also included a small amount of artists and generalizes it into a larger population

# Challenge 3 Submission

Your challenge should be submitted as a **separate file**, **not** at the bottom of your Lab 3 file. Please submit your rendered HTML file. You can copy and paste this code into a **new** Quarto file. Your Challenge 3 submission should only included code necessary for completing the Challenge, nothing else.

You will submit **only** your rendered HTML file. Your HTML file is required to have the following specifications in the YAML options (at the top of your document):

-   be self-contained (`self-contained: true`)
-   include your source code (`code-tools: true`)
-   include all your code and output (`echo: true`)
-   include **no** messages from loading in packages or the data (`messages: false`)

**If any of the options are not included, your Lab 3 or Challenge 3 assignment will receive an "Incomplete" and you will be required to submit a revision.**

